\title{Systems Philosophy - One}
\author{Ryan Spangler}
\date{\today}

\documentclass[11pt]{article}

\usepackage{commath}
\usepackage{graphicx}
\usepackage{listings}
\usepackage{amsfonts}

% python highlighting ----------
\usepackage{color}
\usepackage{listings}
\usepackage{textcomp}
\usepackage{setspace}
%\usepackage{palatino}

%\doublespacing

\setcounter{secnumdepth}{0}

\begin{document}
\maketitle

\section{Interests and Course Expectations}

I am interested in most everything.  This must be part of the reason I am drawn to systems science, I see the connections between things as more significant than the things in and of themselves.  Throughout my life I have endeavored to understand as many different things as I can find, ranging from math to music to neuroscience to programming to quantum mechanics to molecular biology (to name a few).  I have discovered that the more different two things or subjects seem, the greater is the connection between them.  I have learned much about programming from playing music for instance, and vice versa.  This plays into the idea of having multiple perspectives on something as a good way to understand it, where the thing I am taking multiple perspectives on is the totality of existence itself, as seen through the lens of all of these seemingly distinct yet ultimately related fields.  

So if I had to say something I would like to get out of the course, it would be insight into more connections between things that would otherwise seem unrelated.  Beyond that, it would be to hone and expand the ability to perceive connections and make connections that were previously unrecognized.  To learn approaches on how to see connections, and to develop possibly or even glimpse a general theory of connection and relatedness and interactibility.  I want to find some way to reckon with the totality of existence in a way that is not overwhelming or just boggling, but to make the connections and abstractions that reduce the bewildering complexity of all things to something that can be understood and modeled and ultimately influenced.  

This I believe is being served in the sessions we have had so far.  It is hard to talk about *everything* per se, but as much as that is possible, this is what we seem to be doing.

\end{document}  

